---
title: "File System in R"
author: "Joshua Oluoch"
date: "2/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#                          FILE MANAGEMENT IN R

Understanding file management in R is very important for every day user as it dictates how to set environments, input and output data from R, file manipulation among other useful things.

See below some of the quick ways to work with file system in R

### Working Directory 

This is the directory that R system operates from. For projects it defaults to the project folder. To get the current working directory of R, use the command getwd() as below

```{r getwd}
#NOT RUN
#getwd()
```

To set working directory to a folder of your choice, run the command setwd(folder)
```{r setwd}
#NOT RUN
#setwd(/path/to/folder/)
```

### File manipulation
The following functions enable easy handling of files in R.

1. **list.files(directory)** - This function is used to list all the files within a directory.

```{r list_files}
#NOT RUN
#list.files() #This list all the files in the current folder
#To get files in another folder you run
#NOT RUN
#list.files(/path/to/folder/)
```

2. **list.dirs(directory)** - This function list all the folders within the given directory.

```{r list_dirs}
#NOT RUN
#list.dirs() # This list all the folders in the current directory
#To get the folders contained in a different directory, you run list.dirs(/path/to/directory/)
#list.dirs("../") # This lists all the folders of the parent directory, working directory included.
```

3. **file.access(filename) ** - This function tests if the given file exists in the current folder (0-success, 1-failure)
```{r file_access}
file.access("rscfp2019.dta")
```

4. **file.append(x, y)** -This function appends file y to file x, in-situ on the computer.
```{r file_append}
#NOT RUN
#file.append(file_a, file_b)
```

5. **file.choose()** - This function allows for interactive choosing of a file. For instance, if you want to read in a csv file from a different location and you do not want to explicitly write down the file path, you could use file.choose() function.

```{r file_choose}
#NOT RUN
#dat = read.csv(file.choose()) # This helps you navigate to the folder containing the file you want to read

```
6. **OTHER USEFUL FILE MANIPULATION FUNCTIONS**
See other useful file manipulation functions
```{r other_file_functions}
#NOT RUN
#dir.create("tmp") #Creates a directory named 'tmp'

#file.copy("sleep.sav", "tmp") # Copies the file "sleep.sav" to "tmp" folder

#unlink("tmp", recursive = TRUE) #Deletes the folder "tmp"

#file.create(c("a.txt","b.txt")) #creates two files "a.txt" and "b.txt"

#file.remove("a.txt", "b.txt") #Delete two files "a.txt" and "b.txt""

#file.create("a")
#file.rename("a","a_1") #Renames file "a" to "a_1"

```

